Towards a Machine Learning approach for Aggregate Computing

### 2021-11-04 paper-2021-woa-ac-ml.tex
add references, remove old todos

So, we claim that Machine Learning can help us in handling this "unseen" situation, leaving only the burden
 to developers of defining the right collective specifications.
%
This lead to an ``hybrid" approach where a part of behaviour is still expressed using common 
 Aggregate Computing abstractions and another part is distilled through learning~\cite{research}.
%
In this paper, as the first steps towards a so-called \hybridaggregate{}, we investigate how Aggregate Computing can be
 extended with Reinforcement Learning through a case
 study where we outline the performance improvement of the "hop count" program.
\todos{Outline}
\suggestions{I do not know if we can avoid it altogether, but also a brief description of machine learning approaches can be useful in the paper flow}

@@

Aggregate Computing is studied in various simulated scenario (thanks to Alchemist, a meta-simulator for pervasive systems), rangin from smart cities to crodws, thaks to programming language and toolchain that support 
 this paradigm, such as ScaFi, FCCP and Protelis.

@@

@@

\outline{Probably, for a "short" paper, this section can be avoided.}
\outline{walkthrogh traditional engineering way to design CAS}
\outline{walkthrogh Aggregate Computing}
\subsection{Machine Learning}
\suggestions{I do not know, perhaps it is better to avoid this }
\outline{walkthrogh Machine learning approach applied in MARL}
\outline{current limitation (application-specific, not large scale, centralised in some cases,..)}

@@

\todos{present approaches for learning in MAS, e.g., through prominent surveys}

@@

\suggestions{Move the definition of team learning and concurrent learning in the previous section}
\paragraph{Learning problem} 
Multi-agent learning can be divided in \textit{Concurrent} and \textit{Team} learning. 
In the foster, each agent perform the learning process and in the latter,
 the learning process is performed only by one agent and then shared with the collective.
% 
Team learning is usually used in the ``swarm'' system, due to the homogeneous behaviour 
 --- i.e. each agent performs the same program. 
 Here, a common reward signal exists and evaluate the overall collective behaviour.
 Aggregate Computing is quite near to this kind of setting.
% 
Indeed, each agent follows the \textit{same} aggregate program shared within the entire system.

@@

\subsection{Roadmap}
The path towards full integration of Machine Learning and Aggregate Computing is long and probably we do not know the implications that it can bring. 
Currently, we program to:
\begin{enumerate}
  \item Enhance building blocks through learning -- \textit{offline}: 
  in this first step, we need to understand what kind of learning modality is most suitable for us. 
  To decrease the complexity, we aim at perform learning \textit{offline} and at the \textit{building block} level.
  \item Enhance building blocks through learning -- \textit{online}:
  here, the most challenging part in bringing the offline results could be in the reward function definition.
  Indeed, at runtime, we cannot use a global view of the system. 
  A novel way to describe a collective reward by Aggregate Computing might be taken into consideration.
  \item Create building blocks through learning: some building blocks currently lack a proper implementation -- 
  such as the distributed leader election -- so learning can help us also in crafting them. 
  Here, though, the research space increase drastically, 
  so probably it might be not easy to bring our previous results there.
  \item Improve framework aspects through learning: 
  when we successfully apply to learn in the first two steps, 
  we can also consider moving learning at the framework level, to manage non-functional aspects. 
  In this way, developers can focus on the "high level" behaviour aspect.
\end{enumerate}
\todo[inline, color=green]{AC + ML for building block creation -- online}
\todo[inline, color=green]{AC + ML for framework level -- online}

>>>>

So, we claim that Machine Learning can help us in handling this "unseen" situation, leaving only the burden
 to developers of defining the right collective specifications.
%
This lead to an ``hybrid" approach where a part of behaviour is still expressed using common 
 Aggregate Computing abstractions and another part is distilled through learning~\cite{research}.
%
In this paper, as the first steps towards a so-called \hybridaggregate{}, we investigate how Aggregate Computing can be
 extended with Reinforcement Learning through a case
 study where we outline the performance improvement of the "hop count" program.
\todos{Outline}

@@

Aggregate Computing is studied in various simulated scenario (thanks to Alchemist, a meta-simulator for pervasive systems), rangin from smart cities to crodws, thaks to programming language and toolchain that support 
 this paradigm, such as ScaFi~\cite{DBLP:conf/ecoop/CasadeiV16}, FCCP~\cite{DBLP:conf/acsos/Audrito20} and Protelis~\cite{PianiniSAC2015}.

@@

\subsection{ScaFi DSL}

@@

@@

@@

\paragraph{Learning problem} 
Multi-agent learning can be divided in \textit{Concurrent} and \textit{Team} learning. 
In the foster, each agent perform the learning process and in the latter,
 the learning process is performed only by one agent and then shared with the collective.
% 
Team learning is usually used in the ``swarm'' system, due to the homogeneous behaviour 
 --- i.e. each agent performs the same program. 
 Here, a common reward signal exists and evaluate the overall collective behaviour.
 Aggregate Computing is quite near to this kind of setting.
% 
Indeed, each agent follows the \textit{same} aggregate program shared within the entire system.

@@

\subsection{Roadmap}
The path towards full integration of Machine Learning and Aggregate Computing is long and probably we do not know the implications that it can bring. 
Currently, we program to:
\begin{enumerate}
  \item Enhance building blocks through learning -- \textit{offline}: 
  in this first step, we need to understand what kind of learning modality is most suitable for us. 
  To decrease the complexity, we aim at perform learning \textit{offline} and at the \textit{building block} level.
  \item Enhance building blocks through learning -- \textit{online}:
  here, the most challenging part in bringing the offline results could be in the reward function definition.
  Indeed, at runtime, we cannot use a global view of the system. 
  A novel way to describe a collective reward by Aggregate Computing might be taken into consideration.
  \item Create building blocks through learning: some building blocks currently lack a proper implementation -- 
  such as the distributed leader election -- so learning can help us also in crafting them. 
  Here, though, the research space increase drastically, 
  so probably it might be not easy to bring our previous results there.
  \item Improve framework aspects through learning: 
  when we successfully apply to learn in the first two steps, 
  we can also consider moving learning at the framework level, to manage non-functional aspects. 
  In this way, developers can focus on the "high level" behaviour aspect.
\end{enumerate}
